### 2021-02-19 ch06_paper_gpp/paper_gpp.tex
Finished section 6.1

As argued in detail in the previous chapter, single-process-oriented emergent
constraints do not seem to be beneficial in constraining climate projections
from the vastly complex modern-day \acp{ESM} since they might overly simplify
the problem. To overcome this issue, we explore an alternative approach that
expands the \acf{MDER} method with a supervised \acf{ML} algorithm. In contrast
to emergent constraints, this approach considers (1) multiple process-based
predictors, (2) multi-dimensional (gridded) target variables and (3) non-linear
relationships between the predictors and the target variable. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, \ac{ECS} as a single scalar metric for each climate model is not an
appropriate target variable. As an alternative, we analyze projections of
future \acf{GPP} in this study, which constitutes the largest flux of the
terrestrial carbon uptake. As described in detail in
\cref{subsec:02:carbon_cycle_perturbations}, the global carbon cycle actively
slows down global warming by removing \ac{CO2} from the atmosphere. In
particular, the terrestrial biosphere absorbs about $30 \unit{\%}$ of the total
anthropogenic \ac{CO2} emission \autocite{Friedlingstein2020}. Thus, accurately
quantifying the future evolution of the terrestrial carbon sink is crucial for
reliable climate change projections.

@@

This work is already published in \textcite{Schlund2020}. For this paper, the
author of this thesis lead the writing and the analysis and implemented the
code to reproduce this study with all figures and tables using the
\ac{ESMValTool}. \Cref{sec:06:data_and_methods} of this chapter provides an
overview of the data and methods used in this paper. The results are presented
in \cref{sec:06:results} and \cref{sec:06:summary_and_discussion} closes with a
summary and discussion.

@@

In this section, we present the observational emergent constraint on the
\ac{CO2} fertilization effect by \textcite{Wenzel2016} which we use to globally
rescale the models for their individual bias in the sensitivity of \ac{GPP} to
an increase in atmospheric \ac{CO2} concentration. Our objective here is to
present a general methodology and suggest further improvements upon this
constraint that could be used in the future. \textcite{Wenzel2016} show a
correlation between the increase in the \ac{CO2} seasonal cycle amplitude
simulated by models and the projected \ac{CO2} fertilization. When combined
with the observed trends in the \ac{CO2} seasonal cycle amplitude, the
fractional \ac{GPP} change at the time of \ac{CO2} doubling relative to
pre-industrial conditions in a \ac{1BGC} is constrained. The predictor in this
study is the sensitivity of the \ac{CO2} seasonal cycle amplitude to rising
atmospheric \ac{CO2} concentrations, defined as the slope of the linear
regression between the \ac{CO2} seasonal cycle amplitude and the annual mean
atmospheric \ac{CO2} concentrations (see \cref{fig:app:b:co2} for details). The
emergent constraint is physically motivated by the hypothesis that increasing
terrestrial \ac{GPP} is a main driver for the observed changes in the \ac{CO2}
seasonal cycle amplitude \autocite{Gray2014, Keeling1996, Zhao2014}. This
hypothesis is based on the fact that the seasonal cycle in atmospheric \ac{CO2}
concentration originates from photosynthesis and decomposition processes: in
summer, photosynthesis removes \ac{CO2} from the atmosphere; in winter,
additional \ac{CO2} is added to the atmosphere due to the decomposition of
organic matter \autocite{Keeling1995}. Thus, \ac{CO2} fertilization leads to an
increase of the \acs{CO2} seasonal cycle amplitude due to increased \acs{CO2}
removal from the atmosphere. We note that this emergent constraint implicitly
assumes that there is no temperature-driven increase in respiration over time.
Therefore, this approach most likely overestimates the true sensitivity of
\ac{GPP} to an increase in the atmospheric \ac{CO2} concentration.

@@

Since the global constraint above cannot be used directly as a predictor in the
\ac{ML} approach which operates on gridded values (see
\cref{subsec:06:step_2}), we use it as first step to globally rescale the
\ac{ESM} output. This important first step allows us to constrain the large
uncertainty in the global \ac{CO2} fertilization effect of the \acs{CMIP}5
models. In this study, two different gridded target variables are examined (see
\cref{fig:06:schematic_steps}): the monthly climatology of future absolute
\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario (see step 2a) and the fractional \acs{GPP} change over
the \nth{21} century as defined in \cref{eq:06:f} (see step 2b). The global
rescaling of the gridded target variables works as follows: Let $f_m$ be the
global mean fractional \acs{GPP} change over the \nth{21} century of \ac{ESM}
$m$, $f'$ the constrained value of the global mean fractional \ac{GPP} change
and $y_{mi}$ the target variable of climate model $m$. Then, we define the
rescaled target variable $y'_{mi}$ by
\begin{equation}
  y'_{mi} = y_{mi} \cdot \frac{f'}{f_m},
  \label{eq:06:y}
\end{equation}
where the index $i$ counts over all grid cells (and months when absolute
\ac{GPP} is used as target variable in step 2a) and $m$ refers to the \ac{ESM}
($m \in \left\{ 1, 2, \ldots, 7 \right\}$). Thus, values for the individual
grid cells/months are rescaled by a factor that is determined by the ratio
between the models' global mean \ac{GPP} change and the constrained global mean
\ac{GPP} change derived from the \textcite{Wenzel2016} emergent constraint
using the observed change in the \ac{CO2} cycle amplitude. Models that simulate
a too high global mean \ac{GPP} change are nudged toward smaller values and
vice versa. The rescaled target variables are then fed into the \ac{ML} model
(step 2) presented in the following section. We note that this global rescaling
assumes that the relative biases in the fractional \ac{GPP} change of the
individual \acp{ESM} are constant over the whole globe.

@@

\begin{table}[t]
  \centering
  \begin{tabular}[t]{p{0.27\columnwidth} l p{0.11\columnwidth}
      p{0.3\columnwidth}}
    \toprule
    Predictor & \makecell{Observation‐driven \\ data} & Used time range &
    Physical connection to \acs{GPP} \\
    \midrule
    \predictor{GPP} & \makecell{FLUXNET-MTE \\ \autocite{Jung2011}} &
    \range{1991}{2000} & -- \\
    \midrule
    \predictor{LAI} & \makecell{LAI3g \\ \autocite{Zhu2013}} &
    \range{1982}{2005} & \acs{LAI} is a measure for the number of leaves in a
    grid cell. The photosynthesis rate is highly dependent on the number of
    leaves (and vegetation in general). \\
    \midrule
    \predictor{PR} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & Water is essential for the chemical processes of
    photosynthesis. \\
    \midrule
    \predictor{RSDS} & \makecell{ERA-Interim \\ \autocite{Dee2011}} &
    \range{1979}{2005} & Solar radiation is essential for the chemical
    processes of photosynthesis. \\
    \midrule
    \predictor{T} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & \acs{T} and photosynthesis rate have a common driver
    (incoming solar radiation). \\
    \bottomrule
  \end{tabular}
  \caption{Process-oriented diagnostics (\emph{predictors} or \emph{features})
    used in the \acf{GBRT} model to predict the target variables. For step 2a
    (target variable: absolute \acs{GPP}), all listed variables are monthly
    climatologies of the specified time ranges in the historical climate. For
    step 2b (target variable: fractional \acs{GPP} change), the temporal mean
    over the specified time ranges is calculated for all variables.
    \AdaptedFrom{Schlund2020}.}
  \label{tab:06:predictors}
\end{table}

For step 2 we use the \ac{GBRT} algorithm to find an optimal regression
function $F$ relating the target variable $y$ to a set of $K$ predictors
$\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)} \right)$ (\emph{features} or
\emph{covariates}) such that the predicted target variable is given by $\hat{y}
= F(\bm{x})$ \autocite{Death2007, Elith2008, Friedman2001, Friedman2002}. Its
basic elements are decision trees, which use binary splits of the input data to
create decision rules. Due to their simple nature, decision trees are easy to
interpret but cannot be used to create satisfying predictions for very complex
datasets. This issue can be overcome by a technique called \emph{boosting}
\autocite{Freund1996}. Boosting iteratively improves the performance of
\emph{weak learners} (in our case decision trees) in the following way: At
first, a single decision tree is fitted to a randomly selected subsample of the
training data $\left\{ \left( \bm{x}_i, y_i \right) \mid i \in I \right\}$ with
an arbitrary index set $I$ (using a predefined loss function, \eg{} the mean
square error). Then, further decision trees are linearly added to the
regression function $F$ (scaled by a regularization parameter $\nu \ll 1$
called \emph{learning rate}) and the model is fitted again in the same manner.
This ensures that the prediction for poorly modeled training points in the
beginning will gradually improve during the training. The whole procedure is
repeated until the loss function does not improve anymore, as evaluated on an
independent validation data set. Relevant hyperparameters of this algorithm
(maximum number of decision trees, maximum depth of the decision trees,
learning rate and the subsample ratio) are optimized by $5$-fold \ac{CV} using
the training data. More details on the \ac{GBRT} algorithm are given in
\cref{subsec:app:b:gbrt}.

@@

>>>>

As argued in detail in the previous \namecref{ch:05:paper_ecs},
single-process-oriented emergent constraints do not seem to be beneficial in
constraining climate projections from the vastly complex modern-day \acp{ESM}
since they might overly simplify the problem. To overcome this issue, we
explore an alternative approach that expands the \acf{MDER} method (see
\cref{subsec:02:mder}) with a supervised \acf{ML} algorithm. In contrast to
emergent constraints, this approach considers (1) multiple process-based
predictors, (2) multi-dimensional (gridded) target variables and (3) non-linear
relationships between the predictors and the target variable. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, \ac{ECS} as a single scalar metric for each climate model is not an
appropriate target variable. As an alternative, we analyze projections of
future \acf{GPP} in this study, which constitutes the largest flux of the
terrestrial carbon uptake. As described in detail in
\cref{subsec:02:carbon_cycle_perturbations}, the global carbon cycle actively
slows down global warming by removing \ac{CO2} from the atmosphere. In
particular, the terrestrial biosphere absorbs about $30 \unit{\%}$ of the total
anthropogenic \ac{CO2} emission \autocite{Friedlingstein2020}. Thus, accurately
quantifying the future evolution of the terrestrial carbon sink is crucial for
reliable climate change projections.

@@

This work is already published in \textcite{Schlund2020}. For this paper, the
author of this thesis lead the writing and the analysis and implemented the
code to reproduce this study with all figures and tables using the
\ac{ESMValTool}. \Cref{sec:06:data_and_methods} of this
\namecref{ch:06:paper_gpp} provides an overview of the data and methods used in
this paper. The results are presented in \cref{sec:06:results} and
\cref{sec:06:summary_and_discussion} closes with a summary and discussion.

@@

In this \namecref{subsec:06:step_1}, we present the observational emergent
constraint on the \ac{CO2} fertilization effect by \textcite{Wenzel2016} which
we use to globally rescale the models for their individual bias in the
sensitivity of \ac{GPP} to an increase in atmospheric \ac{CO2} concentration.
Our objective here is to present a general methodology and suggest further
improvements upon this constraint that could be used in the future.
\textcite{Wenzel2016} show a correlation between the increase in the \ac{CO2}
seasonal cycle amplitude simulated by models and the projected \ac{CO2}
fertilization. When combined with the observed trends in the \ac{CO2} seasonal
cycle amplitude, the fractional \ac{GPP} change at the time of \ac{CO2}
doubling relative to pre-industrial conditions in a \ac{1BGC} is constrained.
The predictor in this study is the sensitivity of the \ac{CO2} seasonal cycle
amplitude to rising atmospheric \ac{CO2} concentrations, defined as the slope
of the linear regression between the \ac{CO2} seasonal cycle amplitude and the
annual mean atmospheric \ac{CO2} concentrations (see \cref{fig:app:b:co2} for
details). The emergent constraint is physically motivated by the hypothesis
that increasing terrestrial \ac{GPP} is a main driver for the observed changes
in the \ac{CO2} seasonal cycle amplitude \autocite{Gray2014, Keeling1996,
  Zhao2014}. This hypothesis is based on the fact that the seasonal cycle in
atmospheric \ac{CO2} concentration originates from photosynthesis and
decomposition processes: in summer, photosynthesis removes \ac{CO2} from the
atmosphere; in winter, additional \ac{CO2} is added to the atmosphere due to
the decomposition of organic matter \autocite{Keeling1995}. Thus, \ac{CO2}
fertilization leads to an increase of the \acs{CO2} seasonal cycle amplitude
due to increased \acs{CO2} removal from the atmosphere. We note that this
emergent constraint implicitly assumes that there is no temperature-driven
increase in respiration over time. Therefore, this approach most likely
overestimates the true sensitivity of \ac{GPP} to an increase in the
atmospheric \ac{CO2} concentration.

@@

Since the global constraint above cannot be used directly as a predictor in the
\ac{ML} approach which operates on gridded values (see
\cref{subsec:06:step_2}), we use it as first step to globally rescale the
\ac{ESM} output. This important first step allows us to constrain the large
uncertainty in the global \ac{CO2} fertilization effect of the \acs{CMIP}5
models. In this study, two different gridded target variables are examined (see
\cref{fig:06:schematic_steps}): the monthly climatology of future absolute
\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario (see step 2a) and the fractional \acs{GPP} change over
the \nth{21} century as defined in \cref{eq:06:f} (see step 2b). The global
rescaling of the gridded target variables works as follows: Let $f_m$ be the
global mean fractional \acs{GPP} change over the \nth{21} century of \ac{ESM}
$m$, $f'$ the constrained value of the global mean fractional \ac{GPP} change
and $y_{mi}$ the target variable of climate model $m$. Then, we define the
rescaled target variable $y'_{mi}$ by
\begin{equation}
  y'_{mi} = y_{mi} \cdot \frac{f'}{f_m},
  \label{eq:06:y}
\end{equation}
where the index $i$ counts over all grid cells (and months when absolute
\ac{GPP} is used as target variable in step 2a) and $m$ refers to the \ac{ESM}
($m \in \left\{ 1, 2, \ldots, 7 \right\}$). Thus, values for the individual
grid cells/months are rescaled by a factor that is determined by the ratio
between the models' global mean \ac{GPP} change and the constrained global mean
\ac{GPP} change derived from the \textcite{Wenzel2016} emergent constraint
using the observed change in the \ac{CO2} cycle amplitude. Models that simulate
a too high global mean \ac{GPP} change are nudged toward smaller values and
vice versa. The rescaled target variables are then fed into the \ac{ML} model
(step 2) presented in the following \namecref{subsec:06:step_2}. We note that
this global rescaling assumes that the relative biases in the fractional
\ac{GPP} change of the individual \acp{ESM} are constant over the whole globe.

For step 2 we use the \ac{GBRT} algorithm to find an optimal regression
function $F$ relating the target variable $y$ to a set of $K$ predictors
$\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)} \right)$ (also called
\emph{features} or \emph{covariates}) such that the predicted target variable
is given by $\hat{y} = F(\bm{x})$ \autocite{Death2007, Elith2008, Friedman2001,
  Friedman2002}. Its basic elements are decision trees, which use binary splits
of the input data to create decision rules. Due to their simple nature,
decision trees are easy to interpret but cannot be used to create satisfying
predictions for very complex datasets. This issue can be overcome by a
technique called \emph{boosting} \autocite{Freund1996}. Boosting iteratively
improves the performance of \emph{weak learners} (in our case decision trees)
in the following way: At first, a single decision tree is fitted to a randomly
selected subsample of the training data $\left\{ \left( \bm{x}_i, y_i \right)
\mid i \in I \right\}$ with an arbitrary index set $I$ (using a predefined loss
function, \eg{} the mean square error). Then, further decision trees are
linearly added to the regression function $F$ (scaled by a regularization
parameter $\nu \ll 1$ called \emph{learning rate}) and the model is fitted
again in the same manner. This ensures that the prediction for poorly modeled
training points in the beginning will gradually improve during the training.
The whole procedure is repeated until the loss function does not improve
anymore, as evaluated on an independent validation data set. Relevant
hyperparameters of this algorithm (maximum number of decision trees, maximum
depth of the decision trees, learning rate and the subsample ratio) are
optimized by $5$-fold \acf{CV} using the training data. More details on the
\ac{GBRT} algorithm are given in \cref{subsec:app:b:gbrt}.

A convenient way to evaluate and interpret \ac{GBRT} models (or other
statistical models in general) is to identify the most relevant predictors.
Globally, this so-called \emph{feature importance} is closely related to the
number of appearances of a predictor in the different decision trees that set
up the \ac{GBRT} model \autocite{Friedman2001}. To evaluate the feature
importance regionally (for the different grid cells), we use a model
distillation technique called \emph{\ac{LIME}}, which approximates the
\ac{GBRT} model with a local linear regression model for each grid cell
\autocite{Ribeiro2016}. The corresponding linear coefficients $b_k$ for each
predictor $x^{(k)}$ constitute a good proxy to the model's Jacobian
\begin{equation}
  b_k \approx \frac{\partial y}{\partial x^{(k)}},
  \label{eq:06:jacobian}
\end{equation}
which is difficult to compute explicitly for this particular \ac{ML} model.
Thus, features with high (low) absolute $b_k$ have a high (low) impact on the
local prediction of the target variable $y$.

@@

\begin{table}[t]
  \centering
  \begin{tabular}[t]{p{0.27\columnwidth} l p{0.11\columnwidth}
      p{0.3\columnwidth}}
    \toprule
    Predictor & \makecell{Observation‐driven \\ data} & Used time range &
    Physical connection to \acs{GPP} \\
    \midrule
    \predictor{GPP} & \makecell{FLUXNET-MTE \\ \autocite{Jung2011}} &
    \range{1991}{2000} & -- \\
    \midrule
    \predictor{LAI} & \makecell{LAI3g \\ \autocite{Zhu2013}} &
    \range{1982}{2005} & \acs{LAI} is a measure for the number of leaves in a
    grid cell. The photosynthesis rate is highly dependent on the number of
    leaves (and vegetation in general). \\
    \midrule
    \predictor{PR} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & Water is essential for the chemical processes of
    photosynthesis. \\
    \midrule
    \predictor{RSDS} & \makecell{ERA-Interim \\ \autocite{Dee2011}} &
    \range{1979}{2005} & Solar radiation is essential for the chemical
    processes of photosynthesis. \\
    \midrule
    \predictor{T} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & \acs{T} and photosynthesis rate have a common driver
    (incoming solar radiation). \\
    \bottomrule
  \end{tabular}
  \caption{Process-oriented diagnostics (also called \emph{predictors} or
    \emph{features}) used in the \acf{GBRT} model to predict the target
    variables. For step 2a (target variable: absolute \acs{GPP}), all listed
    variables are monthly climatologies of the specified time ranges in the
    historical climate. For step 2b (target variable: fractional \acs{GPP}
    change), the temporal mean over the specified time ranges is calculated
    for all variables. \AdaptedFrom{Schlund2020}.}
  \label{tab:06:predictors}
\end{table}

As already described in the previous \namecref{subsec:06:step_1}, we analyze
two different target variables in this study: the monthly climatology of
absolute \ac{GPP} at the end of the \nth{21} century (step 2a) and the
fractional \ac{GPP} change over the \nth{21} century (step 2b). A schematic
representation of step 2a is shown in \cref{fig:06:mlr_concept}. The target
variable $y$ for the \ac{ML} model is the rescaled monthly climatology of
future GPP at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario given by \cref{eq:06:y}. We relate this target variable
to several process-oriented diagnostics $\bm{x}$ of the past climate
(historical simulation) listed in \cref{tab:06:predictors}. A single training
point $\left( \bm{x}_i, y_i \right)$ for the supervised \ac{ML} algorithm
corresponds to the set of the aforementioned diagnostics evaluated on a single
grid cell/month of a single climate model (see colored points in
\namecrefs{fig:06:mlr_concept} \labelcref{fig:06:mlr_concept}{\color{BrownRed}
  a} and \labelcref{fig:06:mlr_concept}{\color{BrownRed} b}). Step 2b differs
from step 2a only in the target variable used. Both steps are independent
approaches and are evaluated separately.

The main part of steps 2a and 2b each consist of two phases. In the first phase
(training of the algorithm), we build a statistical model based on the
empirical relationship between observable process-oriented variables of the
present-day climate and future \ac{GPP} in the \acs{CMIP}5 ensemble (see gray
surface in \namecrefs{fig:06:mlr_concept}
\labelcref{fig:06:mlr_concept}{\color{BrownRed} a} and
\labelcref{fig:06:mlr_concept}{\color{BrownRed} b}). This has a flavor similar
to the \ac{MDER} algorithm \commentcite{Karpechko2013, Senftleben2020,
  Wenzel2016a}{see \cref{subsec:02:mder}}. However, in contrast to the original
\ac{MDER} algorithm, our multivariate \ac{GBRT} approach is able to handle
multiple predictors with non-linear dependencies. Moreover, instead of
considering a single globally or regionally averaged value per climate model,
our approach uses spatiotemporally gridded data for every \ac{ESM} in the
ensemble. This dramatically increases the available number of training data
(even if they are not statistically independent) and uniquely enables the
statistical model to identify and exploit regional processes. Furthermore, the
large training database allows us to perform extensive out-of-sample testing,
which is crucial to ensure the \ac{ML} model's robustness.

In the second phase (predicting with the trained \ac{GBRT} model), we use
observational data products to constrain the \acs{RCP}8.5 \acs{CMIP}5
projection of \ac{GPP}. To do this, observational data for all predictors
$\bm{x}$ of the historical climate (for every grid cell/month) are fed into the
trained \ac{GBRT} model in order to get a spatially constrained projection of
\ac{GPP} (see black points in \cref{fig:06:mlr_concept}{\color{BrownRed} b}).
Similar to the \ac{MDER} algorithm, our approach assumes that there exists an
inter-model relationship between the predictors and the target variable which
also holds for the true climate. This assumption may seem weak at first glance,
but compared to the assumption of traditional weighting approaches (models that
are better in simulating the present-day climate are also better in simulating
the future climate and vice versa), the \ac{MDER} assumption is more reasonable
and allows correcting models instead of just weighing
\commentcite{Karpechko2013}{see \cref{subsec:02:mder}}. Following other
\ac{MDER} studies \autocite{Karpechko2013, Senftleben2020, Wenzel2016a}, we
also select diagnostics as predictor variables $\bm{x}$ that are related to
\ac{GPP}-relevant processes (see \cref{tab:06:predictors} for detailed physical
explanations of the links between the predictors and \ac{GPP}).

The raw \acs{CMIP}5 and observation-driven datasets are preprocessed before
entering the \ac{ML} algorithm (see \cref{subsec:app:b:data_preprocessing}). In
particular, increases greater than $300 \unit{\%}$ in the fractional \ac{GPP}
change that is used as target variable in step 2b are removed. These
unrealistically high values (making up about $5.9 \unit{\%}$ of the data) occur
in places with very small absolute \ac{GPP} values in the historical simulation
that are used in the denominator of the derivation of the fractional \ac{GPP}
change (see \cref{eq:06:f}). In addition, we randomly select $25 \unit{\%}$ of
the training data and use it as hold-out test data. This dataset is excluded
from the training and can therefore be used for an independent validation of
our statistical model.

@@

The estimation of the \acf{SPE} for the \ac{GBRT} models needs to consider
three sources of uncertainty: the uncertainty of the \ac{ML} model itself, the
uncertainty in the rescaling of the target variable (see \cref{eq:06:y}) and
the uncertainty in the input data, that is, errors in the observation-based
products. First, the \ac{SPE} of the \ac{GBRT} model itself (\ie{} uncertainty
in the position of the gray surface in \namecrefs{fig:06:mlr_concept}
\labelcref{fig:06:mlr_concept}{\color{BrownRed} a} and
\labelcref{fig:06:mlr_concept}{\color{BrownRed} b}) is estimated as the
\emph{\acf{RMSEP}}, that is, the \ac{RMSE} between the predicted and true
values of the independent hold-out test dataset \autocite{Bishop2006}. This
approach is justified when the distribution of residuals is unbiased (see
\cref{subsec:app:b:residuals}), in which case the \ac{RMSEP} equals the
standard deviation of the residuals (which we interpret as error).
\Cref{fig:app:b:residuals} shows that this is approximately correct.

The \ac{SPE} of the \ac{GBRT} model can be understood as combined error due to
the internal variability of the climate system, the different model responses
within the climate model ensemble to forcing and the incomplete description of
the climate system with the limited number of variables used as input features.
Second, the uncertainty in the rescaling is simply assessed as the uncertainty
given by the emergent constraint from step 1. Finally, the uncertainty in the
observation-based data (\ie{} uncertainty in the position of the black points
in \cref{fig:06:mlr_concept}{\color{BrownRed} b}) is assessed by error
propagation. The necessary sensitivities of the target variable to the various
predictors for each grid cell/month are given by the linear coefficients $b_k$
given by the \ac{LIME} approach introduced in \cref{subsec:06:step_2}. All
sources of uncertainty are assumed independent and are combined by adding the
squared errors, yielding a single estimate of the \ac{SPE} for each grid
cell/month. Further details on the calculation of errors are given in
\cref{subsec:app:b:uncertainty}.

The predictive power of the \ac{GBRT} algorithm and its robustness is assessed
using a \emph{leave-one-model-out \ac{CV}} approach \commentcite{Elia2002,
  Karpechko2013}{see \cref{fig:06:mlr_concept}{\color{BrownRed} c} and
  \cref{subsec:02:model_weighting}}. For this, a single climate model is removed
from the multi-model ensemble and considered as being the true climate. Our
statistical model is then fitted on all remaining climate models and a
prediction for the \enquote{true} model is created. This allows a simple
evaluation of the predictive power by calculating the \ac{RMSEP}. The whole
process is repeated for every climate model of the ensemble in order to get a
distribution of \acp{RMSEP}. Since this approach is not limited to \ac{GBRT},
it allows a simple comparison of the predictive skill of different statistical
models. We compare the \ac{GBRT} model to the simple \acs{CMIP}5 \acf{MMM} and
a \ac{LASSO} model \autocite{Tibshirani1996}. The \ac{LASSO} model is an
extended linear regression model that regularizes the least squares solution
with an L$1$ penalty term. In contrast to the ordinary least squares
regression, the \ac{LASSO} model promotes sparse solutions (\ie{} only a few
non-zero coefficients/weights), which acts as an intrinsic feature selection
and provides an effective way to combat overfitting induced by collinearity.
